---
title: "Final Project Report - Iris dataset"
author: "Xiaohan Li"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r, set-chunk-opts, echo = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



# Introduction

## Description of the data set

 - This data set contains 150 observations, with 5 variables: sepal length, sepal width, petal length, petal width, and species. 
 - Each observation represents an iris plant.
 - The species has 3 categories: Setosa, Versicolor, and Virginica, with 50 observations for each species.
 - One species is linearly separable from the other 2, the latter are not linearly separable from each other.


## Analysis objectives

 - Use data visualization to explore the relationships between variables.
 - Use data visualization to explore the differences between 3 species.

# Read in data
```{r, load-data}
here::i_am("iris_report.Rmd")

data <- readRDS(
  file = here::here("data/data_clean.rds"))
```


# Tabular Analysis

Characteristics of the `r nrow(data)` plants from the iris data set are displayed in the table below.

```{r, table}
table <- readRDS(
  file = here::here("output/table.rds")
)
table
```

 - From this table, we can get a general overview of the differences among the 3 species in sepal and petal measurements, including length and width.

 - For __setosa__ species, the average sepal length, petal length, and petal width are all below the overall average, while its sepal width is above the average.

 - Species __setosa__ has the smallest average sepal length, petal length, and petal width, but the largest sepal width.
 
 - Species __versicolor__ has the smallest average sepal width.
 
 - Species __virginica__ has the largest average sepal length, petal length and petal width.


# Graphical Analysis
## Boxplot that showing the sepal length differences between 3 species
```{r, boxplot_sepal_length}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/boxplot_sepal_length.png")
)
```

From this boxplot, we can directly see the differences in sepal length among 3 species. 

 - The average sepal length order is setosa < versicolor < virginica.
 - A taller box indicates larger variability. The variability of sepal length in __setosa__ is smaller than that of __versicolor__ and __virginica__.


## Boxplot that showing the sepal width differences between 3 species
```{r, boxplot_sepal_width}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/boxplot_sepal_width.png")
)
```

From this boxplot, we can directly see the differences in sepal width among 3 species. 

 - The average sepal length order is versicolor < virginica < setosa.
 
 - A taller box indicates larger variability. The variability of sepal width in __virginica__ is smaller than that of __versicolor__ and __setosa__.